Active Damping of Rotating Positioning Platforms using Force Feedback

### 2020-07-01 paper/paper.tex
Shrinked two figures to make them subfigures

% Created 2020-07-01 mer. 10:02
% Intended LaTeX compiler: pdflatex
\documentclass{ISMA_USD2020}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{siunitx}
\usepackage{amsmath,amssymb,amsfonts, cases}
\usepackage{algorithmic, graphicx, textcomp}
\usepackage{xcolor, import, hyperref}
\usepackage{subcaption}
\usepackage[USenglish, english]{babel}
\setcounter{footnote}{1}
\input{config.tex}
\author[1,3] {T. Dehaeze}
\author[1,2] {C. Collette}
\affil[1] {Precision Mechatronics Laboratory\NewLineAffil University of Liege, Belgium \NewAffil}
\affil[2] {BEAMS Department\NewLineAffil Free University of Brussels, Belgium \NewAffil}
\affil[3] {European Synchrotron Radiation Facility \NewLineAffil Grenoble, France e-mail: \textbf{thomas.dehaeze@esrf.fr}}
\bibliographystyle{IEEEtran}
\usepackage{tikz}
\usetikzlibrary{shapes.misc,arrows,arrows.meta}
\date{}
\title{Active Damping of Rotating Positioning Platforms using Force Feedback}
\hypersetup{
 pdfauthor={},
 pdftitle={Active Damping of Rotating Positioning Platforms using Force Feedback},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.0.91 (Org mode 9.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org38db2d2}
\label{sec:introduction}
Due to gyroscopic effects, the guaranteed robustness properties of Integral Force Feedback do not hold.
Either the control architecture can be slightly modified or mechanical changes in the system can be performed.
This paper has been published
The Matlab code that was use to obtain the results are available in \cite{dehaeze20_activ_dampin_rotat_posit_platf}.

\section{Dynamics of Rotating Positioning Platforms}
\label{sec:orgd0419a7}
\label{sec:dynamics}
\subsection{Model of a Rotating Positioning Platform}
\label{sec:org4c3a41a}
In order to study how the rotation of a positioning platforms does affect the use of force feedback, a simple model of an X-Y positioning stage on top of a rotating stage is developed.

@@

\subsection{Equations of Motion}
\label{sec:orga2d4956}
To obtain of equation of motion for the system represented in Figure \ref{fig:system}, the Lagrangian equations are used:
\begin{equation}
\label{eq:lagrangian_equations}
  \frac{d}{dt} \left( \frac{\partial L}{\partial \dot{q}_i} \right) + \frac{\partial D}{\partial \dot{q}_i} - \frac{\partial L}{\partial q_i} = Q_i
\end{equation}
with \(L = T - V\) the Lagrangian, \(D\) the dissipation function, and \(Q_i\) the generalized force associated with the generalized variable \(\begin{bmatrix}q_1 & q_2\end{bmatrix} = \begin{bmatrix}d_u & d_v\end{bmatrix}\).

@@

\subsection{Transfer Functions in the Laplace domain}
\label{sec:org16be790}
To study the dynamics of the system, the differential equations of motions \eqref{eq:eom_coupled} are transformed in the Laplace domain and the transfer function matrix from \(\begin{bmatrix}F_u & F_v\end{bmatrix}\) to \(\begin{bmatrix}d_u & d_v\end{bmatrix}\) is obtained:
\begin{equation}
\label{eq:Gd_mimo_tf}
  \begin{bmatrix} d_u \\ d_v \end{bmatrix} = \bm{G}_d \begin{bmatrix} F_u \\ F_v \end{bmatrix}
\end{equation}
with \(\bm{G}_d\) a \(2 \times 2\) transfer function matrix
\begin{equation}
\label{eq:Gd_m_k_c}
  \bm{G}_{d} =
  \begin{bmatrix}
    \frac{ms^2 + cs + k - m \Omega^2}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2} & \frac{2 m \Omega s}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2} \\
    \frac{-2 m \Omega s}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2} & \frac{ms^2 + cs + k - m \Omega^2}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2}
  \end{bmatrix}
\end{equation}

@@

\subsection{System Dynamics and Campbell Diagram}
\label{sec:org0ed334c}
The poles of \(\bm{G}_d\) are the complex solutions \(p\) of
\begin{equation}
  \left( \frac{p^2}{{\omega_0}^2} + 2 \xi \frac{p}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{p}{\omega_0} \right)^2 = 0
\end{equation}

@@

\section{Decentralized Integral Force Feedback}
\label{sec:org284335d}
\label{sec:iff}
\subsection{Force Sensors and Control Architecture}
\label{sec:org5b8858f}
In order to apply Decentralized Integral Force Feedback to the system, force sensors are added in series with the two actuators (Figure \ref{fig:system_iff}).
Two identical controllers \(K_F\) are added to feedback each of the sensed forces to its collocated actuator.
The control diagram is shown in Figure \ref{fig:control_diagram_iff}.

@@

\subsection{Plant Dynamics}
\label{sec:orge8dea8f}
The forces measured by the force sensors are equal to:
\begin{equation}
\label{eq:measured_force}
  \begin{bmatrix} f_{u} \\ f_{v} \end{bmatrix} =
  \begin{bmatrix} F_u \\ F_v \end{bmatrix} - (c s + k)
  \begin{bmatrix} d_u \\ d_v \end{bmatrix}
\end{equation}

@@

\subsection{Decentralized Integral Force Feedback with Pure Integrators}
\label{sec:org8280bcd}
The two IFF controllers \(K_F\) are pure integrators
\begin{equation}
  \bm{K}_F(s) = \begin{bmatrix} K_F(s) & 0 \\ 0 & K_F(s) \end{bmatrix}, \quad K_F(s) = g \cdot \frac{1}{s}
\end{equation}
where \(g\) is a scalar value representing the gain of the controller.

@@

\section{Integral Force Feedback with High Pass Filters}
\label{sec:org6cc7c03}
\label{sec:iff_hpf}
\subsection{Modification of the Control Low}
\label{sec:org9575a34}
\begin{equation}
\label{eq:IFF_LHF}
  K_{F}(s) = g \cdot \frac{1}{s} \cdot \underbrace{\frac{s/\omega_i}{1 + s/\omega_i}}_{\text{HPF}} = g \cdot \frac{1}{s + \omega_i}
\end{equation}

\subsection{Feedback Analysis}
\label{sec:orgbc7c7f2}
\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/loop_gain_modified_iff.pdf}
\caption{\label{fig:loop_gain_modified_iff}Bode Plot of the Loop Gain for IFF with and without the HPF with \(\omega_i = 0.1 \omega_0\), \(g = 2\) and \(\Omega = 0.1 \omega_0\)}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/root_locus_modified_iff.pdf}
\caption{\label{fig:root_locus_modified_iff}Root Locus for IFF with and without the HPF, \(\Omega = 0.1 \omega_0\)}
\end{figure}

\subsection{Optimal Control Parameters}
\label{sec:org6e31c47}

@@

\section{Integral Force Feedback with Parallel Springs}
\label{sec:org8681b34}
\label{sec:iff_kp}
\subsection{Stiffness in Parallel with the Force Sensor}
\label{sec:org0dff3a7}
Stiffness can be added in parallel to the force sensor to counteract the negative stiffness due to centrifugal forces.
If the added stiffness is higher than the maximum negative stiffness, then the poles of the IFF damped system will stay in the (stable) right half-plane.

@@

\subsection{Plant Dynamics}
\label{sec:orgc00a18e}

@@

\subsection{Effect of the Parallel Stiffness on the Plant Dynamics}
\label{sec:orgdbd366f}
\begin{equation}
  \begin{aligned}
    \alpha > \frac{\Omega^2}{{\omega_0}^2} \\
    \Leftrightarrow k_p > m \Omega^2
  \end{aligned}
\end{equation}

@@

\subsection{Optimal Parallel Stiffness}
\label{sec:org8a2fab9}
\begin{figure}[htbp]
\begin{subfigure}[c]{0.49\linewidth}
\includegraphics[width=\linewidth]{figs/root_locus_iff_kps.pdf}
\caption{\label{fig:root_locus_iff_kps} Three values of \(k_p\)}
\end{subfigure}
\begin{subfigure}[c]{0.49\linewidth}
\includegraphics[width=\linewidth]{figs/root_locus_opt_gain_iff_kp.pdf}
\caption{\label{fig:root_locus_opt_gain_iff_kp} \(k_p = 5 m \Omega^2\), optimal damping is shown}
\end{subfigure}
\caption{\label{fig:root_locus_iff_kps_opt}Root Locus for IFF when parallel stiffness is used, \(\Omega = 0.1 \omega_0\)}
\centering
\end{figure}

\section{Comparison of the Proposed Active Damping Techniques for Rotating Positioning Stages}
\label{sec:orgf0c901c}
\label{sec:comparison}
\subsection{Physical Comparison}
\label{sec:orgb81fedd}

\subsection{Attainable Damping}
\label{sec:org9975e71}

@@

\subsection{Transmissibility and Compliance}
\label{sec:org0a87dc5}
\begin{figure}[htbp]
\begin{subfigure}[c]{0.45\linewidth}
\includegraphics[width=\linewidth]{figs/comp_compliance.pdf}
\caption{\label{fig:comp_compliance} Transmissibility}
\end{subfigure}
\begin{subfigure}[c]{0.45\linewidth}
\includegraphics[width=\linewidth]{figs/comp_transmissibility.pdf}
\caption{\label{fig:comp_transmissibility} Compliance}
\end{subfigure}
\caption{\label{fig:comp_active_damping}Comparison of the two proposed Active Damping Techniques, \(\Omega = 0.1 \omega_0\)}
\centering
\end{figure}

\section{Conclusion}
\label{sec:orgcf2c965}
\label{sec:conclusion}

\section*{Acknowledgment}
\label{sec:orga6bcde3}

>>>>

% Created 2020-07-01 mer. 10:48
% Intended LaTeX compiler: pdflatex
\documentclass{ISMA_USD2020}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{siunitx}
\usepackage{amsmath,amssymb,amsfonts, cases}
\usepackage{algorithmic, graphicx, textcomp}
\usepackage{xcolor, import, hyperref}
\usepackage{subcaption}
\usepackage[USenglish, english]{babel}
\setcounter{footnote}{1}
\input{config.tex}
\author[1,3] {T. Dehaeze}
\author[1,2] {C. Collette}
\affil[1] {Precision Mechatronics Laboratory\NewLineAffil University of Liege, Belgium \NewAffil}
\affil[2] {BEAMS Department\NewLineAffil Free University of Brussels, Belgium \NewAffil}
\affil[3] {European Synchrotron Radiation Facility \NewLineAffil Grenoble, France e-mail: \textbf{thomas.dehaeze@esrf.fr}}
\bibliographystyle{IEEEtran}
\usepackage{tikz}
\usetikzlibrary{shapes.misc,arrows,arrows.meta}
\date{}
\title{Active Damping of Rotating Positioning Platforms using Force Feedback}
\hypersetup{
 pdfauthor={},
 pdftitle={Active Damping of Rotating Positioning Platforms using Force Feedback},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.0.91 (Org mode 9.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org6a18e43}
\label{sec:introduction}
Due to gyroscopic effects, the guaranteed robustness properties of Integral Force Feedback do not hold.
Either the control architecture can be slightly modified or mechanical changes in the system can be performed.
This paper has been published
The Matlab code that was use to obtain the results are available in \cite{dehaeze20_activ_dampin_rotat_posit_platf}.

\section{Dynamics of Rotating Positioning Platforms}
\label{sec:orgd72985f}
\label{sec:dynamics}
\subsection{Model of a Rotating Positioning Platform}
\label{sec:org617200e}
In order to study how the rotation of a positioning platforms does affect the use of force feedback, a simple model of an X-Y positioning stage on top of a rotating stage is developed.

@@

\subsection{Equations of Motion}
\label{sec:orgb84906d}
To obtain of equation of motion for the system represented in Figure \ref{fig:system}, the Lagrangian equations are used:
\begin{equation}
\label{eq:lagrangian_equations}
  \frac{d}{dt} \left( \frac{\partial L}{\partial \dot{q}_i} \right) + \frac{\partial D}{\partial \dot{q}_i} - \frac{\partial L}{\partial q_i} = Q_i
\end{equation}
with \(L = T - V\) the Lagrangian, \(D\) the dissipation function, and \(Q_i\) the generalized force associated with the generalized variable \(\begin{bmatrix}q_1 & q_2\end{bmatrix} = \begin{bmatrix}d_u & d_v\end{bmatrix}\).

@@

\subsection{Transfer Functions in the Laplace domain}
\label{sec:orgda6662b}
To study the dynamics of the system, the differential equations of motions \eqref{eq:eom_coupled} are transformed in the Laplace domain and the transfer function matrix from \(\begin{bmatrix}F_u & F_v\end{bmatrix}\) to \(\begin{bmatrix}d_u & d_v\end{bmatrix}\) is obtained:
\begin{equation}
\label{eq:Gd_mimo_tf}
  \begin{bmatrix} d_u \\ d_v \end{bmatrix} = \bm{G}_d \begin{bmatrix} F_u \\ F_v \end{bmatrix}
\end{equation}
with \(\bm{G}_d\) a \(2 \times 2\) transfer function matrix
\begin{equation}
\label{eq:Gd_m_k_c}
  \bm{G}_{d} =
  \begin{bmatrix}
    \frac{ms^2 + cs + k - m \Omega^2}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2} & \frac{2 m \Omega s}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2} \\
    \frac{-2 m \Omega s}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2} & \frac{ms^2 + cs + k - m \Omega^2}{\left( m s^2 + cs + k - m \Omega^2 \right)^2 + \left( 2 m \Omega s \right)^2}
  \end{bmatrix}
\end{equation}

@@

\subsection{System Dynamics and Campbell Diagram}
\label{sec:org006f494}
The poles of \(\bm{G}_d\) are the complex solutions \(p\) of
\begin{equation}
  \left( \frac{p^2}{{\omega_0}^2} + 2 \xi \frac{p}{\omega_0} + 1 - \frac{{\Omega}^2}{{\omega_0}^2} \right)^2 + \left( 2 \frac{\Omega}{\omega_0} \frac{p}{\omega_0} \right)^2 = 0
\end{equation}

@@

\section{Decentralized Integral Force Feedback}
\label{sec:org586ab0c}
\label{sec:iff}
\subsection{Force Sensors and Control Architecture}
\label{sec:org56ae682}
In order to apply Decentralized Integral Force Feedback to the system, force sensors are added in series with the two actuators (Figure \ref{fig:system_iff}).
Two identical controllers \(K_F\) are added to feedback each of the sensed forces to its collocated actuator.
The control diagram is shown in Figure \ref{fig:control_diagram_iff}.

@@

\subsection{Plant Dynamics}
\label{sec:org6b43274}
The forces measured by the force sensors are equal to:
\begin{equation}
\label{eq:measured_force}
  \begin{bmatrix} f_{u} \\ f_{v} \end{bmatrix} =
  \begin{bmatrix} F_u \\ F_v \end{bmatrix} - (c s + k)
  \begin{bmatrix} d_u \\ d_v \end{bmatrix}
\end{equation}

@@

\subsection{Decentralized Integral Force Feedback with Pure Integrators}
\label{sec:orgaedf7bc}
The two IFF controllers \(K_F\) are pure integrators
\begin{equation}
  \bm{K}_F(s) = \begin{bmatrix} K_F(s) & 0 \\ 0 & K_F(s) \end{bmatrix}, \quad K_F(s) = g \cdot \frac{1}{s}
\end{equation}
where \(g\) is a scalar value representing the gain of the controller.

@@

\section{Integral Force Feedback with High Pass Filters}
\label{sec:org34f8977}
\label{sec:iff_hpf}
\subsection{Modification of the Control Low}
\label{sec:org809db54}
In order to limit the low frequency loop gain, an high pass filter (HPF) can be added to the controller.
The controller becomes
\begin{equation}
\label{eq:IFF_LHF}
  \bm{K}_F(s) = \begin{bmatrix} K_F(s) & 0 \\ 0 & K_F(s) \end{bmatrix}, \quad K_{F}(s) = g \cdot \frac{1}{s} \cdot \underbrace{\frac{s/\omega_i}{1 + s/\omega_i}}_{\text{HPF}} = g \cdot \frac{1}{s + \omega_i}
\end{equation}
This is equivalent as to slightly shifting to pole to the left along the real axis.
This modification of the IFF controller is typically done to avoid saturation associated with the pure integrator \cite{preumont91_activ}.

\subsection{Feedback Analysis}
\label{sec:orga2d434f}
\begin{minipage}[c]{0.45\linewidth}
\begin{center}
\includegraphics[scale=1]{figs/loop_gain_modified_iff.pdf}
\captionof{figure}{\label{fig:loop_gain_modified_iff}Bode Plot of the loop gain for IFF with and without the HPF with, \(g = 2\), \(\omega_i = 0.1 \omega_0\) and \(\Omega = 0.1 \omega_0\)}
\end{center}
\end{minipage}
\hfill
\begin{minipage}[c]{0.5\linewidth}
\begin{center}
\includegraphics[scale=1]{figs/root_locus_modified_iff.pdf}
\captionof{figure}{\label{fig:root_locus_modified_iff}Root Locus for IFF with and without the HPF, \(\omega_i = 0.1 \omega_0\) and \(\Omega = 0.1 \omega_0\)}
\end{center}
\end{minipage}

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/loop_gain_modified_iff.pdf}
\caption{\label{fig:loop_gain_modified_iff}Bode Plot of the Loop Gain for IFF with and without the HPF with \(\omega_i = 0.1 \omega_0\), \(g = 2\) and \(\Omega = 0.1 \omega_0\)}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[scale=1]{figs/root_locus_modified_iff.pdf}
\label{fig:root_locus_modified_iff}
\end{figure}

\subsection{Optimal Control Parameters}
\label{sec:org7cc60ca}

@@

\section{Integral Force Feedback with Parallel Springs}
\label{sec:org8ffeef1}
\label{sec:iff_kp}
\subsection{Stiffness in Parallel with the Force Sensor}
\label{sec:org831f255}
Stiffness can be added in parallel to the force sensor to counteract the negative stiffness due to centrifugal forces.
If the added stiffness is higher than the maximum negative stiffness, then the poles of the IFF damped system will stay in the (stable) right half-plane.

@@

\subsection{Plant Dynamics}
\label{sec:org98721cc}

@@

\subsection{Effect of the Parallel Stiffness on the Plant Dynamics}
\label{sec:org0b03ba2}
\begin{equation}
  \begin{aligned}
    \alpha > \frac{\Omega^2}{{\omega_0}^2} \\
    \Leftrightarrow k_p > m \Omega^2
  \end{aligned}
\end{equation}

@@

\subsection{Optimal Parallel Stiffness}
\label{sec:org0aa60d8}
\begin{figure}[htbp]
\begin{subfigure}[c]{0.49\linewidth}
\includegraphics[width=\linewidth]{figs/root_locus_iff_kps.pdf}
\caption{\label{fig:root_locus_iff_kps} Three values of \(k_p\)}
\end{subfigure}
\begin{subfigure}[c]{0.49\linewidth}
\includegraphics[width=\linewidth]{figs/root_locus_opt_gain_iff_kp.pdf}
\caption{\label{fig:root_locus_opt_gain_iff_kp} \(k_p = 5 m \Omega^2\), optimal damping is shown}
\end{subfigure}
\caption{\label{fig:root_locus_iff_kps_opt}Root Locus for IFF when parallel stiffness is used, \(\Omega = 0.1 \omega_0\)}
\centering
\end{figure}

\section{Comparison of the Proposed Active Damping Techniques for Rotating Positioning Stages}
\label{sec:orgde2e2f0}
\label{sec:comparison}
\subsection{Physical Comparison}
\label{sec:org661a4d0}

\subsection{Attainable Damping}
\label{sec:orgb54beda}

@@

\subsection{Transmissibility and Compliance}
\label{sec:org63e90d2}
\begin{figure}[htbp]
\begin{subfigure}[c]{0.45\linewidth}
\includegraphics[width=\linewidth]{figs/comp_compliance.pdf}
\caption{\label{fig:comp_compliance} Transmissibility}
\end{subfigure}
\begin{subfigure}[c]{0.45\linewidth}
\includegraphics[width=\linewidth]{figs/comp_transmissibility.pdf}
\caption{\label{fig:comp_transmissibility} Compliance}
\end{subfigure}
\caption{\label{fig:comp_active_damping}Comparison of the two proposed Active Damping Techniques, \(\Omega = 0.1 \omega_0\)}
\centering
\end{figure}

\section{Conclusion}
\label{sec:org646f615}
\label{sec:conclusion}

\section*{Acknowledgment}
\label{sec:org7bb2645}
